---
title: Spotify Data Analysis
author: ~
date: '2020-12-05'
slug: spotify-data-analysis
categories: []
tags: []
---

```{r, message=FALSE}
library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)
spotify <- read_csv(here::here("data", "tidytuesday",
                               "data","2020","2020-01-21","spotify_songs.csv"))
```

* Question 1: How does energy compare to the duration of a song? To test this I will create a chart, possibly a point or violin, to show dispersion and look for clustering at certain energy levels and times to see if there is a trend for how long high energy songs typically are and how long low energy songs are.
```{r}
spotify %>%
  ggplot(aes(x=duration_ms, y = energy)) +
  geom_point() +
  theme_solarized() +
  xlab("Song Duration in Milliseconds") +
  ylab("Song Energy") +
  ggtitle("Song Energy vs Song Duration in Milliseconds")
```

* Question 2: What are the 10 most popular songs? What genres show up more than once? To test this I will need to filter for genre, song name, and popularity and then arrange them in descending order. From there I will get a list of the top 10 popular songs and can see what genres show up more than once.
```{r}
spotify %>%
  select(track_name, playlist_genre, track_popularity) %>%
  group_by(track_name) %>%
  arrange(desc(track_popularity))
```

* Question 3: How does speechiness compare to popularity? Is there a sweetspot? To test this I will need to graph speechiness against popularity and look for a trend, I will likely start with a point graph and change from there if there is a better solution that presents itself.
```{r}
spotify %>%
  ggplot(aes(x = speechiness, y = track_popularity)) +
  geom_point() +
  theme_solarized() + 
  labs(x="Speechiness",y="Song Popularity") +
  ggtitle("Speechiness vs Song Popularity")
```
Seems to be a strange trend where it slopes downward from top popularity to around .4 speechiness and slopes upward from .12 popularity to .4 speechiness to form a triangular point. The ones at the bottom people hated for reasons other than speechiness. 











